---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# nofrills

_Low-Cost Anonymous Functions_

[![Travis-CI Build Status](https://travis-ci.org/egnha/nofrills.svg?branch=master)](https://travis-ci.org/egnha/nofrills)
[![codecov](https://codecov.io/gh/egnha/nofrills/branch/master/graph/badge.svg)](https://codecov.io/gh/egnha/nofrills)

## Overview

_nofrills_ is a tiny R package that provides a function, `fn()`, that enables 
you to create (anonymous) functions, of _arbitrary_ call signature, 
economically. It is a drop-in replacement for the usual `function(<arguments>)
<body>` invocation, but costs less:

  - It is **shorter**:
    ```R
    fn(x, y = 1 ~ x + y)
    
    ..(x, y = 1 ~ x + y)
    ```
    are both equivalent to
    ```R
    function(x, y = 1) x + y
    ```
    
  - It is **safer**: by enabling
    [quasiquotation](http://rlang.tidyverse.org/reference/quasiquotation.html),
   `fn()` allows you to “burn in” values, which guards your function from being
   affected by unexpected scope changes (see example, below).


## Installation

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("egnha/nofrills")
```

## Usage

```{r, echo = FALSE}
library(nofrills)
```

### Same syntax as `function()` but shorter

```{r}
fn(x ~ x + 1)

fn(x, y ~ x + y)

fn(x, y = 2 ~ x + y)

fn(x, y = 1, ... ~ log(x + y, ...))

# the only exception, cf. alist()
fn(x, ... = , y ~ log(x + y, ...))

fn(~ NA)
```

### Supports quasiquotation

#### Unquoting values

```{r}
z <- 0

fn(x, y = !! z ~ x + y)

fn(x ~ x > !! z)
```

#### Unquoting argument names

```{r}
arg <- "y"

fn(x, !! arg := 0 ~ x + !! as.name(arg))
```

#### Splicing in argument lists

```{r}
args <- alist(x, y = 0)

fn(!!! args, ~ x + y)  # note the one-sided formula
```

### Protect functions against scope changes

Both `f()` and `f_solid()` return the same value of x

```{r}
x <- "x"

f <- function() x
f_solid <- fn(~ !! x)

f()

f_solid()
```

But if the binding `x` is (unwittingly) changed, `f()` changes, while
`f_solid()` remains unaffected.

```{r}
x <- sin

f()

f_solid()
```

### 😃 functions

Pop quiz: Both of these smileys produce functions.

```{r, eval = FALSE}
..(~8^D)
..(8~D)
```

But which one is actually callable?

## Alternatives

The following packages provide alternative anonymous-function constructors. 
Unlike `fn()`, they automatically configure the resulting function’s call
signature, so they can afford to be more concise.

  - [lambda](https://github.com/jimhester/lambda) provides `f()`. It uses a
    `bquote()`-like notation for function declaration, which, by forgoing 
    explicit call signature specification, is very compact. However,
    quasiquotation is not supported, as it wasn't available when lambda was
    developed.

  - [rlang](https://github.com/tidyverse/rlang) provides 
    [`as_function()`](http://rlang.tidyverse.org/reference/as_function.html), 
    which allows you to create anonymous functions of an arbitrary number of
    arguments, named `..1` (`.x`), `..2` (`.y`), `..3`, etc. The call signature
    is fixed.
    
## Acknowledgement

The [rlang](https://github.com/tidyverse/rlang) package by [Lionel
Henry](https://github.com/lionel-) and [Hadley 
Wickham](https://github.com/hadley) makes _nofrills_ possible. Crucially, rlang
provides the engine for quasiquotation and expression capture.

## License

MIT Copyright © 2017 [Eugene Ha](https://github.com/egnha)
